### 2025-04-07 sports-cognitive-psychology.qmd
20250407 - Lochbaum meta-analysis

The Yerkes–Dodson law describes the association between arousal and performance, as depicted in @fig-yerkesDodsonLaw.
For simple tasks, performance tends to increase with arousal.
For difficult tasks, there is an inverted-U-shaped association between arousal and performance, as depicted in @fig-hebbianYerkesDodsonLaw.
For difficult tasks, too little arousal is bad because it leads to insufficient attention on the task at hand; however, too much arousal is also bad because anxiety can impair performance through its negative effects on attention (e.g., "tunnel vision"), memory, decision making, problem-solving, and multitasking.
That is, there is an optimal range of arousal for complex and challenging tasks, such as those in sports, which often require divided attention and multitasking.

>>>>

Meta-analyses have demonstrated that there are a number of psychological factors that influence player performance in sports.
For instance, a meta analysis by @Lochbaum2022 found that psychological processes such as confidence and mindfulness were associated with better performance in sports.

One process that may have a nonlinear association with performance is arousal or anxiety.
The Yerkes–Dodson law describes the association between arousal and performance, as depicted in @fig-yerkesDodsonLaw.
For simple tasks, performance tends to increase with arousal.
For difficult tasks, there is an inverted-U-shaped association between arousal and performance, as depicted in @fig-hebbianYerkesDodsonLaw.
For difficult tasks, too little arousal is bad because it leads to insufficient attention on the task at hand; however, too much arousal is also bad because anxiety can impair performance through its negative effects on attention (e.g., "tunnel vision"), memory, decision making, problem-solving, and multitasking.
That is, there is an optimal range of arousal for complex and challenging tasks, such as those in sports, which often require divided attention and multitasking.
